---
title: "dc.synth.dmg"
author: "Aaron Schroeder"
date: "11/10/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## A Data Commons Package for Generating Block and Parcel-Level Demographics using ACS Estimates and Counts of Housing Units per Census Block

### Installation

```{r eval=FALSE}
remotes::install_github("uva-bi-sdad/dc.synth.dmg")
```

```{r message=FALSE}
library(dc.synth.dmg)
```

### Get ACS and Block Address Count Data

#### get_data() returns a list of 2 datasets, one of ACS data, the other a count of total housing units per Census block

```{r message=FALSE, warning=FALSE, comment=NA}
mydata <- get_data(state_abbrev = "VA", county_fips = "013")

print(mydata)
```

### Generate Demographic Estimates per Census Block

```{r message=FALSE, comment=NA}
bk_dmgs <-
  generate_block_dmgs(acs_data = mydata$acs_data_51013,
                      bac_data = mydata$bac_data_51013)

print(head(bk_dmgs[order(geoid)]))
```

### To change the default ACS variables set your own named list of variables

```{r message=FALSE, comment=NA, warning=FALSE}
set_acs_variables(list(my_total_population = "B01001_001",
                       my_male_population = "B01001_002"))

mydata <- get_data(state_abbrev = "VA", county_fips = "013")

bk_dmgs <-
  generate_block_dmgs(acs_data = mydata$acs_data_51013,
                      bac_data = mydata$bac_data_51013)

print(head(bk_dmgs[order(geoid)]))
```
